A Streaming Algorithm for 2-Center with Outliers in High Dimensions

### 2015-05-03 paper.tex
Second submission
Query added

To obtain our result, we have used a combination of several ideas
including parallelization construction, far/close ball separation,
and keeping a lower bound on the radius and distance of the optimal balls.
We have also utilized ideas used in~\cite{kim2014improved} 
for the 2-center problem with no outliers.
%However, our algorithm is much detailed and uses many new ingredients.
However, our problem is much harder here, % handling outliers makes the problem much harder,
as we not only need to find balls of minimum radius, 
but we also need to decide which subset of points to cluster.
This is in particular more challenging in the streaming model,
where we only have a single pass over the input, and we must decide on the fly
which point is an outlier, 
and which one can be safely ignored as a non-outlier point,
%using only a limited amount of storage.
to comply with the working space restriction enforced by the model.

@@

\begin{algorithm}
\caption{\sc Query} 
\algsetup{indent=1.5em}
\label{alg:query}
\begin{algorithmic}[1]
%\Function{Query}{}
	\STATE solutions $\gets \set{\Call{ComputeMin}{B_1, B_2, \Buffer}}$

	\STATE candidates $\gets \Buffer$
	\IF {$\card{B_2} < \dz$}
		\STATE candidates $\gets \Buffer \cup B_2 \setminus \set{c_2}$
	\ELSE
		\STATE $B_1 \gets B_u$
	\ENDIF

	\STATE $\delta_0 = \delta$
	\FOR{$c_2 \in$ candidates}
		\STATE $\rc  \gets \frac{2}{\alpha} \len{c_1 c_2}$
		\STATE $B'_1 \gets B_1$, $\delta \gets \max \{ \delta_0, r \}$
		\STATE $B'_2 \gets \emptyset$, $\Buffer' \gets \emptyset$ 
		\FOR{$p \in$ candidates}
			\IF {\NOT (\Call{AddToB$'_1$}{$p$} \OR \Call{AddToB$'_2$}{$p$})}
				\STATE add $p$ to $\Buffer'$
			\ENDIF
		\ENDFOR
		\STATE add \Call{ComputeMin}{$B'_1$, $B'_2$, $\Buffer'$} to solutions
	\ENDFOR
	\RETURN $\min\set{\text{solutions}}$
%\EndFunction
\end{algorithmic}
\end{algorithm}

@@

Now, we show that the query algorithm correctly considers all possible candidates for $c_2$.
Let $C$ denotes the set of candidate points for $c_2$.
By invariants (c), we know that $B_1 \cap B_2^* = \emptyset$. 
Therefore, $(B_2 \cup \Buffer) \cap B_2^* \neq \emptyset$,
and hence, $C \subseteq (B_2 \cup \Buffer) \cap B_2^*$. % in general.
We now use the following lemma.
%In this general case, our candidates $C$ for $c_2$ is set $(B_2 \cup \Buffer) \cap B_2^*$.

@@

\noindent
By Lemma \ref{lem:B2dz}, when $|B_2| \geq \dz$, 
we known that $(\{c_2\} \cup \Buffer) \cap B_2^* \neq \emptyset$. 
Therefore, if $|B_2| \ge \dz$, then our candidates for $C$ are $\{c_2\} \cup \Buffer$.

>>>>

To obtain our result, we have used a combination of several ideas
including parallelization construction, far/close ball separation, centerpoint theorem,
and keeping a lower bound on the radius and distance of the optimal balls.
We have also utilized ideas used in~\cite{kim2014improved} 
for the 2-center problem with no outliers.
%However, our algorithm is much detailed and uses many new ingredients.
However, our problem is much harder here, % handling outliers makes the problem much harder,
as we not only need to find balls of minimum radius, 
but we also need to decide which subset of points to cluster.
This is in particular more challenging in the streaming model,
where we only have a single pass over the input, and we must decide on the fly
which point is an outlier, 
and which one can be safely ignored as a non-outlier point,
%using only a limited amount of storage.
to comply with the working space restriction enforced by the model.

@@

\begin{algorithm}
\caption{\sc Query} 
\algsetup{indent=1.5em}
\label{alg:query}
\begin{algorithmic}[1]
	%\STATE solutions $\gets \set{\Call{ComputeMin}{B_1, B_2, \Buffer}}$
	\STATE solutions $\gets \set{}$

	\STATE candidates $\gets \set{c_2} \cup \Buffer $
	\IF {$\card{B_2} < \dz$}
		\STATE add $B_2$ to candidates
	\ELSE
		\STATE $B_1 \gets B_u$
	\ENDIF

	\STATE $\delta_0 = \delta$
	\FOR{$c \in$ candidates}
		\STATE $\rc  \gets \frac{2}{\alpha} \len{c_1 c}$
		\STATE $B'_1 \gets B_1$, $\delta \gets \max \{ \delta_0, r \}$
		\STATE $B'_2 \gets \emptyset$, $\Buffer' \gets \emptyset$ 
		\FOR{$p \in$ candidates}
			\IF {\NOT (\Call{AddToB$'_1$}{$p$} \OR \Call{AddToB$'_2$}{$p$})}
				\STATE add $p$ to $\Buffer'$
			\ENDIF
		\ENDFOR
		\STATE add \Call{ComputeMin}{$B'_1$, $B'_2$, $\Buffer'$} to solutions
	\ENDFOR
	\RETURN $\min\set{\text{solutions}}$
%\EndFunction
\end{algorithmic}
\end{algorithm}

@@

@@

\noindent
%Now, we show that the query algorithm correctly considers all possible candidates for $c_2$.
Let $C$ denote the set of candidates for $c_2$.
By invariant~(c), we know that $B_1 \cap B_2^* = \emptyset$. 
Therefore, there exists a point in $(B_2 \cup \Buffer) \cap B_2^*$,
and hence, $C$ is $(B_2 \cup \Buffer) \cap B_2^*$ in general.
%We now use the following lemma.
%In this general case, our candidates $C$ for $c_2$ is set $(B_2 \cup \Buffer) \cap B_2^*$.
However, when $|B_2| \geq \dz$, 
we know by Lemma \ref{lem:B2dz} that $(\{c_2\} \cup \Buffer) \cap B_2^* \neq \emptyset$. 
Therefore, if $|B_2| \ge \dz$, then our candidates for $C$ are only $\{c_2\} \cup \Buffer$.
